---
title: "MotorTrend"
author: "Willianto"
date: "14/06/2020"
output:
  html_document:
    df_print: paged
---

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```

```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(pander) ##Load package for tidy T test result
```

## Motor Trend Data Analysis

### 1. Executive Summary

Motor Trend, a magazine about the automobile industry is looking at a data set of a collection of cars. The company is interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

Regression models and exploratory data analyses are used to answer the above questions.

The information about the Motor Trend dataset can be found at the following link: [mtcars info](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) <-click here

<br />


### 2. Load the mtcars data and perform some basic exploratory data analyses
***
```{r mtcars}
library(datasets)
data(mtcars) ##Load mtcars dataset
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual')) ##assign automatic and manual transmission type to AM
```
<br />

From the fig. 1 in the appendix we know that the mtcars dataset contains `r nrow(mtcars)` observations and `r length(mtcars)` variables: `r names(mtcars)`. The fig. 2 boxplot shows the relationship between MPG and AM, from the boxplot we can see that the cars with manual transmission yield a better miles per gallon than the car with automatic transmission.

<br />

### 3. Regression Models
***

However we need to test whether it is true that the cars with manual transmission yield better mpg than cars with automatic tranmission by performing a hyphothesis testing (T Test). The null hypothesis is the transmission does not impact the result of MPG and the alternative hyphothesis is the tranmission will impact the result of MPG. To reject null hypothesis, a scientific standard of more than 95% confidence interval is used because anything less is no significance difference for scientific studies. Hence the P value (critical value) must be less than 5% ( .05) for a significance difference. If the P value is less than .05, it is likely that the transmission has impact on the MPG. If the P value is more than .05, it is unlikely that the transmission has impact on MPG. 

```{r hypothesisTest}
##T test to show whether transmission has impact on MPG
hTest <- t.test(mpg~am, data=mtcars, paired=F, var.equal=T, conf.level=0.95) 
hTest$p.value
```
The P value results of the hypothesis T test conducted shows the p value is `r hTest$p.value` which is less than .05. Therefore we reject the null hypothesis and we can conclude that tranmission type has an impact on the MPG. The T Test result of the mtcars dataset also shows that the mean for automatic is 17.15 MPG and the mean for manual is 24.39 MPG.


<br />

### 5. Appendix

***
Figure 1:
```{r dataSummary}
kable(summary(mtcars))
dim(mtcars)
names(mtcars)
```
<br />

Figure 2:
```{r boxplot}
##Exploratory Data Analysis by looking at the relationship of MPG and AM using boxplot
boxplot(mpg~am, data = mtcars,
        xlab = "Transmission",
        ylab = "Miles per Gallon",
        main = "MPG by Transmission Type")
```

Figure 3: 
