---
title: "MotorTrend"
author: "Willianto"
date: "14/06/2020"
output:
  html_document:
    df_print: paged
---

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```

```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(pander) ##Load package for tidy T test result
```

## Motor Trend Data Analysis

### 1. Executive Summary

Motor Trend, a magazine about the automobile industry is looking at a data set of a collection of cars. The company is interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”
"Quantify the MPG difference between automatic and manual transmissions"

The information about the Motor Trend dataset can be found at the following link: [mtcars info](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html) <-click here

<br />


### 2. Load the mtcars data and perform some basic exploratory data analyses
```{r mtcars}
library(datasets)
data(mtcars)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```
<br />

From the fig. 1 in the appendix we know that the mtcars dataset contains `r nrow(mtcars)` observations and `r length(mtcars)` variables: `r names(mtcars)`. The fig. 2 boxplot shows the relationship between MPG and AM, looking from the boxplot we can see that the manual tranmission yield a better MPG than automatic transmission.

<br />


### 5. Appendix

***
Figure 1:
`r kable(summary(mtcars))`

<br />

Figure 2:
```{r boxplot}
boxplot(mpg~am, data = mtcars,
        xlab = "Transmission",
        ylab = "Miles per Gallon",
        main = "MPG by Transmission Type")
```